---
title: "Make a Plot"
author: ""
format:
  html:
    embed-resources: true
editor: visual
execute:
  warning: false
editor_options:
  chunk_output_type: console
---

## Task 1: Import

The required packages for this homework exercise have already been added to the following code-chunk.

1.  Run the following code chunk. Tipp: Click on the green play button in the top right corner of the code chunk.

```{r}
library(gapminder)
library(ggplot2)
library(dplyr)
```

## Task 2: Transform for data in 2007

**Fill in the gaps**

1.  A code chunk has already been created below.

2.  Start with the `gapminder` object and add the pipe operator at the end of the line.

3.  On a new line use the `filter()` function to narrow down the data for observation of the year 2007.

4.  Use the assignment operator to assign the data to an object named `gapminder_2007`.

5.  Run the code contained in the code chunk and fix any errors.

6.  Next to the code chunk option `#| eval:` change the value from `false` to `true`.

7.  Render the document and fix any errors.

```{r}
#| eval: false

gapminder_2007 <- ___ |>
  ___(year == 2007)
```

## Task 3: Summarize data for life expectancy by continent

**Fill in the gaps**

1.  A code chunk has already been created.

2.  Start with the `gapminder_2007` object and add the pipe operator at the end of the line.

3.  On a new line use the `group_by()` function to group the operations that follow by continent. Add the pipe operator at the end of the line.

4.  On a new line use the `summarise()` function to calculate the number of observations (count) and median life expectancy.

5.  Use the assignment operator to assign the data to an object named `gapminder_summary_2007`.

6.  Run the code contained in the code chunk and fix any errors.

7.  Next to the code chunk option `#| eval:` change the value from `false` to `true`.

8.  Render the document and fix any errors.

```{r}
#| eval: false

gapminder_summary_2007 <- ___ |>
  group_by(___) |>
  summarise(
    count = n(),
    lifeExp = median(___)
  )
```

## Task 4: Summarize data for life expectancy by continent and year

**Fill in the gaps**

1.  A code chunk has already been created.

2.  Start with the `gapminder` object and add the pipe operator at the end of the line.

3.  On a new line use the `group_by()` function to group the operations that follow by continent and year. Add the pipe operator at the end of the line.

4.  On a new line use the `summarise()` function to calculate and median life expectancy.

5.  Use the assignment operator to assign the data to an object named `gapminder_summary_continent_year` .

6.  Run the code contained in the code chunk and fix any errors.

7.  Next to the code chunk option `#| eval:` change the value from `false` to `true`.

8.  Render the document and fix any errors.

```{r}
#| eval: false

gapminder_summary_continent_year <- gapminder ___
  group_by(continent, ___) |>
  ___(lifeExp = median(lifeExp))

```

## Task 13: Data communication

**In the YAML header (between the three dashes at the top of the document)**

1.  Add your name as the author of this document
2.  Render the document and fix any errors

## Task 14: Complete assignment

### Stage, Commit & Push to GitHub

1.  Open the Git pane in RStudio. It's in the top right corner in a separate tab.
2.  **Stage** your changes by checking appropriate box next to all files (if you select one file with your mouse, you can then highlight them all with Ctrl + A on your keyboard and check all boxes)
3.  Write a meaningful commit message (e.g. "Completed homework assignment 02.) in the **Commit message** box
4.  Click **Commit**. Note that every commit needs to have a commit message associated with it.

### Open an issue on GitHub

Once you have ensured that the Quarto document renders without errors and you have pushed all your changes to GitHub, you can complete the assignment by opening an issue on

1.  Open [github.com](https://github.com/) in your browser.
2.  Navigate to the GitHub organisation for the course.
3.  Find the repository md-02-assignments that ends with your GitHub username.
4.  Click on the "Issues" tab.
5.  Click on the green "New issue" button.
6.  In the "Title" field write: "Completed module 2 assignments".
7.  In the "Leave a comment" field, tag the course instructors @larnsce and ask some questions.
