---
title: "Make a Plot"
author: ""
format:
  html:
    embed-resources: true
editor: visual
execute:
  warning: false
editor_options:
  chunk_output_type: console
---

## Task 1: Import

The required packages for this homework exercise have already been added to the following code-chunk.

1.  Run the following code chunk. Tipp: Click on the green play button in the top right corner of the code chunk.

```{r}
library(gapminder)
library(ggplot2)
library(dplyr)
```

## Task 2: Transform for data in 2007

**Read and understand the code**

1.  Read the code below out loud, verbally describing what each line does.

2.  Run the code chunk below.

3.  Observe the output in the console.

```{r}
gapminder |>
  filter(year == 2007)
```

4.  Write in your own words what you think this code does:

*Your answer here:*

5.  Now run the code chunk below that assigns the result to an object.

6.  **Tip:** Watch the Environment tab in RStudio (top right pane) to see the new object appear.

```{r}
gapminder_2007 <- gapminder |>
  filter(year == 2007)
```

7.  Click on `gapminder_2007` in the Environment tab to view the data. What do you notice about this dataset compared to the original `gapminder` dataset?

*Your answer here:*

## Task 3: Summarize data for life expectancy by continent

**Read and understand the code**

1.  Read the code below out loud, verbally describing what each line does.

2.  Run the code chunk below.

3.  Observe the output in the console.

```{r}
gapminder_2007 |>
  group_by(continent) |>
  summarise(
    count = n(),
    lifeExp = median(lifeExp)
  )
```

4.  Write in your own words what you think this code does:

*Your answer here:*

5.  Now run the code chunk below that assigns the result to an object.

6.  **Tip:** Watch the Environment tab to see how this new object differs from `gapminder_2007`.

```{r}
gapminder_summary_2007 <- gapminder_2007 |>
  group_by(continent) |>
  summarise(
    count = n(),
    lifeExp = median(lifeExp)
  )
```

7.  Look at the `gapminder_summary_2007` object in the Environment tab. How many rows and columns does it have? Why is this different from `gapminder_2007`?

*Your answer here:*

## Task 4: Summarize data for life expectancy by continent and year

**Read and understand the code**

1.  Read the code below out loud, verbally describing what each line does.

2.  Run the code chunk below.

3.  Observe the output in the console.

```{r}
gapminder |>
  group_by(continent, year) |>
  summarise(lifeExp = median(lifeExp))
```

4.  Write in your own words what you think this code does:

*Your answer here:*

5.  Now run the code chunk below that assigns the result to an object.$

6.  **Tip:** Notice how the Environment tab shows different information for different types of objects.

```{r}
gapminder_summary_continent_year <- gapminder |>
  group_by(continent, year) |>
  summarise(lifeExp = median(lifeExp))
```

7.  Compare this result with Task 3. What's different about grouping by two variables instead of one? How does this affect the output?

*Your answer here:*

## Task 5: Create your first plot

Now that you understand how to transform data, let's create a plot.

1.  Read the code below out loud, verbally describing what you think it will do.

2.  Run the code chunk.

```{r}
ggplot(data = gapminder_2007, 
       mapping = aes(x = continent, 
                     y = lifeExp)) +
  geom_boxplot()
```

3.  Write in your own words what this plot shows:

*Your answer here:*

## Task 6: Data communication

**In the YAML header (between the three dashes at the top of the document)**

1.  Add your name as the author of this document

2.  Render the document and fix any errors

## Task 7: Commit & Push & Issue

-   Refer back to the instructions for Step 2 and Step 3 on the course website to complete the assignment 02a: <https://ds4owd-002.github.io/website/assignments/md-02/am-02-1-data-visualisation.html#step-2-commit-and-push-your-changes>
